### 2020-12-08 ms/ms.tex
Pre-revamp

@@

The natural environment is changing rapidly. Over the last century,
habitats have become increasingly fragmented and
isolated\supercite{Haddad2015}, the climate has become warmer, and
extreme climate events more frequent\supercite{IPCC2014}. Biological
diversity---the species of plants and animals with which we cohabit
the Earth---has responded predictably. Present-day extinction rates
are estimated to be up to three orders of magnitude greater than
background rates \supercite{Pimm2014} and are projected to increase
further over the next several decades \supercite{MEA2005}. Current
estimates suggest that over one million species may be threatened with
extinction\supercite{IPBESland2019}, and that we have experienced a
50\% decline in animal diversity in the past 40 years
\supercite{McLellan2014}.

The main direct human-induced drivers that impact biodiversity now are
habitat loss \supercite{Newbold2015} and fragmentation
\supercite{Haddad2015}, but climate change is likely to become a
dominant driver in the next few
decades\supercite{Parmesan2003,Thuiller2007,Urban2015}. Projected
impacts of climate change on biodiversity have attracted much
attention, but the uncertainty around the magnitude of future
extinctions (\emph{e.g.}, 7.9\%\supercite{Urban2015} vs.\
37\%\supercite{Thomas2004} of species) highlights the inadequacy of
current forecasts.  In the face of changing climate species must
either shift in space, to track favourable conditions, or time, for
example, flowering and breeding earlier.
% I've dropped the migration reference here because that's shifting in
% both space and time, and so sort of ruins the either/or structure
Evidence suggests they are doing both
\supercite{Parmesan2003,Parmesan2006,Menzel2006,Chen2011}, but not all
species are responding equally. Some species are expanding their
distributions, sometimes invasively \supercite{Hellmann2008}, and some
species are flowering later\supercite{Cook2012}; species may also be
simply becoming less predictable \supercite{Pearse2017phenology},
perhaps as evolved responses to environmental cues break down.  Simply
quantifying species' mean responses ignores the important variability
in responses among species \supercite{Wolkovich2012}. Developing
robust predictions of ecological responses to climate change thus
requires data from diverse species, ecosystems, and climates to
account for this large inter-specific variation.

The velocity of temperature change---the rate a species would have to
shift its range to maintain constant temperature---has been calculated
at an average of 0.42 km/yr \supercite{Loarie2009}. A meta-analysis of
taxonomic groups estimated the average pole-ward shift in species'
distributions has been approximately 1.69 km/yr \supercite{Chen2011},
suggesting that species might be capable of keeping pace with shifting
climate. Yet even this estimate is derived from just 21 studies with
limited taxonomic coverage (361 birds, 9 mammals, and no plants), and
critically any single point estimate does not reflect the large
variation in responses among species or within taxa. Here we leverage
the vast volume of distributional data stored in digital data
repositories\supercite{url_gbif}, representing over 10,000 species of
plants, fungi, mammals, birds, reptiles, amphibians, and insects, to
provide the first global synthesis of species distributional responses
to climate change. By contrasting historical niche dimensions with
those of the present day, we can characterise the realised biotic
velocity of climate change, which represents the relative degree to
which species track climate across space.

We show that, on average, species' historical (pre-1980) ranges have
warmed significantly, and that species have generally shifted their
distributions to ameliorate this warming, but not sufficiently to keep
pace with climate change. Yet this general pattern masks a remarkably
degree of variability in species responses, and many species' ranges
are now even warmer than if their distributions had remained
static. Further, we find that, while species' niche dimensions are
strongly predicted by their evolutionary history and functional
traits, the same is not true of their degree of tracking. Thus while
species' distributions are reasonably predictable, their changes in
distributions are not.

@@

Where $current_\tau$ is a given quantile ($\tau$) of a species'
climatic distribution across its present-day range, $past_\tau$ the
same quantile of a species' range in the past, and $projected_\tau$
the same quantile of a species' past range under present-day climatic
conditions. This index should be calculated separately for
temperature, pollution, or whatever environmental variable is of
interest. Thus a $track$ of 0 would indicate a species that has
tracked changing climate perfectly, while a $track$ of 1 would be
consistent with a species that had not moved whatsoever to track
changing environmental conditions. Values between 0 and 1 suggest
moderate tracking, while values greater than 1 or less than 0
represent a change in climate relationship greater than if the species
had not moved at all. Uncertainty in either species' past or current
ranges can be accounted for through a bootstrap procedure, whereby the
past and current distributions are re-sampled to standardise sampling
across time periods (see methods).

Our approach has two main innovations: its focus on (1) the whole of a
species' distribution and (2) expected change. (1) When estimating
species' range distributions we typically focus on species'
limits\supercite{Sexton2009} (\emph{e.g.}, upper and lower quantiles)
while in community ecological \supercite{Tilman2004} and
macro-evolutionary \supercite{Zanne2018} studies we tend to focus on
central tendencies (\emph{e.g.}, the median). However, there is no
strong \emph{a priori} reason that species can, or must, track all
quantiles equally. Indeed some evidence suggests different constraints
on species' warm-range limits versus their cool-range limits, with
biotic interactions more influential in the former, and climate
determining the latter\supercite{Freeman2018}.  It is possible,
therefore, that the drivers of species' range edges and centres may be
diverging under climate change. Our approach is to check all quantiles
of a species' distribution. (2) The magnitude of experienced and
projected climate change shows substantial geographical variation
\supercite{Loarie2009}, and so species' responses might thus be
expected to also vary in space. We suggest that species'
distributional responses should be measured on the basis of the
magnitude of change to which they are exposed, and the denominator of
our index re-scales species' change around an expectation under
distributional stasis.  This change reflects the magnitude of
ecological and evolutionary selection, and while there are many ways
of addressing this, we suggest ours is perhaps the simplest and most
transparent, so represents a fair starting point. We note that we
could also rescale our index according to the distance that would need
to be travelled to maintain climate stasis (\emph{sensu}
\citeauthor{Loarie2009}\supercite{Loarie2009}), but prefer to present
here a unitless index of tracking, which is simpler to compare across
taxa and between biogeographic regions.

\textbf{Niche tracking across the tree of life throughout the past
  sixty years.} To assess the extent to which species have tracked
climate, we used the most comprehensively sampled time-series data for
species' distributions (GBIF\supercite{url_gbif}) and climate (UEA's
CRU\supercite{Harris2014}). This represents over 630 million raw
occurrence records collected over the past century. We focus on five
quantiles: the \nth{5}, \nth{25}, \nth{50} (median), \nth{75}, and
\nth{95}. Here, we present species' range data collected from
1955--2015 for 10,700 species (4,879 plants, 1,517 fungi, 273 mammals,
433 birds, 147 reptiles and amphibians, and 3,451 insects), each
observation matched to one of nine climate variables in the year in
which they were sampled (mean, minimum, maximum temperature,
precipitation, `rainy-day' frequency, vapour pressure, potential
evapotranspiration, `frost day' frequency, and cloud cover). In the
methods we give more detail of our reproducible analysis pipeline, and
in the supplementary materials provide instructions for repeating our
analyses with different data curation and cleaning choices. Our metric
was designed to account for known biases in the GBIF data
\supercite{Beck2014}, and in the methods we outline our simulations
which address the influence of sampling uncertainty.

Almost all species are tracking some climate variable through time
(99.5\% have at least one $track_\tau$ between 0 and 1). Most species
are, to some extent, tracking temperature (see Figure \ref{overview});
65\% track at least one quantile and 43\% track at least two.
% Giving an index in units of distance is a good idea; I think it
% needs some thought, though, because our index is unitless because we
% divide the units through; I'm not sure the number we get out of this
% would be directly interpretable like theirs.
The three temperature indices are, however, the least-tracked of any
climate variable, while precipitation is the most-tracked (86\% of
species track at least one of its quantiles). Notably, however, many
taxa are not tracking climate through time, and many have experienced
more change than if their distributions had remained static. Just as
most species are tracking at least one quantile of one climate
variable, 99.9\% of species have `overshoot' (have a $track_\tau$ less
than 0; for example becoming colder in a warming region) in at least
one quantile of at least one climate variable. As we show in Figure
\ref{overview}, these results are consistent with our null simulations
where species' range change is a combination of essentially random
movement and a moderate degree of subsequent climate filtering. Thus
while our results confirm that, by-and-large, species are tracking
climate change, the magnitude of this varies across species, taxa, and
aspects of climate (see Figure \ref{overview}). For example, while
fungi and insects are, on the whole, broadly tracking the median of
temperature (median $track_{50\%}$ of 0.46 and 0.64, respectively),
plants and mammals track temperature (median $track_{50\%}$ of 1.16
and 1.07, respectively) more poorly than precipitation (median
$track_{5\%}$ of 0.79 and 0.89, respectively).

\textbf{Idiosyncratic drivers of species' tracking.} To examine the
potential drivers of species' ranges and range tracking, we first
examined the phylogenetic signal of species' climate quantiles. We
focused on mammals, birds, reptiles, and plants, since we were able to
find broadly inclusive, dated phylogenies for these taxa. We found
strong evidence of phylogenetic signal in species' past (median
Pagel's $\lambda$\supercite{Pagel1999} of 0.60) and current (median
$\lambda$ of 0.57) distributions, but limited signal in our index
(median $\lambda$ of $<$0.001). Past climate showed stronger
phylogenetic signal than our index in 158 of all 180 (4 taxa, 5
quantiles, and 9 climate variables) possible taxon-climate-quantile
combinations (144 for current climate; see Figures \ref{phylo} and
\ref{traits-signal}). This confirms the known result that species'
current-day and past ranges are strongly influenced by
evolution\supercite{Wiens2010}, but suggests that species' realised
velocity of change (our tracking index) may be associated with
different factors.

We additionally examined species' traits as potential drivers of
species' distributions and potential to track changing environment. We
focused on mammals, birds, and plants, since global, openly-released
trait datasets exist for these taxa. Again, we found statistically
significant associations between traits and past distributions of 191
of 270 (at $\alpha_{5\%}$; 6 traits, 9 climate variables, and 5
quantiles) taxon-climate-quantile combinations (193 for current
distributions), but not between the traits and our tracking index (13
combinations or 5\%; see Figure \ref{traits-signal}). While the traits
we used are strongly associated with species' climate
responses\supercite{Wright2004,McCain2014} (hence their strong
predictive power of species' distributions, but not degree of
tracking), we emphasise that we could not find sufficient data to
assess the potential impact of dispersal
ability\supercite{Schloss2012}. These findings, while surprising, are
in keeping with emerging evidence that species' traits are a
relatively poor predictor of change at range edges
\supercite{Angert2011}.

\textbf{Species' movement through climate space is disconnected from
  their position within it.} Given our index is poorly predicted by
traditionally important variables, we now consider whether index
values are predictive of each other. It is well known that the axes of
global climate are not independent and that they are not changing
independently\supercite{Harris2014,IPCC2014}, and so we would expect
species' relative tracking of climate to show similar patterns. As we
show in Figure \ref{pca}, principal components analysis of species'
current and past climate distributions, which we refer to as their
climate space, shows strong correlations across climate variables. Yet
our index does not; species' relative degree of tracking is both of
much higher dimension (the amount of variance explained by each
principal component axis is similar) and shows relatively less
correlation across climates axes . We argue, therefore, that while
species' positions in climate space show strong associations, species'
relative tracking of their position in climate space does not. This
result explains the poor predictive power of species' traits and
evolutionary history for our index: the ecological and evolutionary
rules that determine species' climate relationships are
changing. Thus, in our analysis, we can explain species' current
distributions only by the degree to which they resemble past
distributions.

There are alternative explanations for our results, of which we
highlight three: (1) we are ignoring important niche axes, (2) we are
ignoring biotic interactions, and (3) our index highlights random
variation in distribution. (1) Our climate variables are not
all-encompassing, and we ignore other drivers such as geology and
habitat-type. Yet we do detect strong pattern in current and past
distributions, suggesting that the absence of pattern in our indices
is meaningful. (2) Competition and facilitation among species are
likely strong drivers of species' ability to track
climate\supercite{Davis1999}. Yet, again, they are surely also likely
to have driven species' past distributions, and so it is unclear to us
why only our metric, and not also current and past distributions, is
so poorly predicted by our trait and phylogenetic data. (3) If species
were perfectly tracking climate, $current_\tau$ would be equal to
$past_\tau$ and so our index values could reflect sampling error or
some other purely stochastic input. Yet such noise, if truly random,
would mean our empirical index values would be centred at 0 or
resemble our simulated data, neither or which is the case (see Figure
\ref{overview}).

\clearpage
\subsection*{Conclusion}
Our results are consistent with the processes underlying species'
responses to climate change being distinct from those that generated
their distributions in the past. This is in keeping with emerging
evidence from the paleo-ecological literature that species' historic
ranges were more variable than previously thought, and seemingly
influenced by different environmental factors than today
\supercite{Veloz2012,Maguire2015}. During and after the last ice-age,
for example, `no-analogue' conditions generated previously-unseen
species assemblages \supercite{Williams2007}. Equally, our observation
of idiosyncrasy in species' degree of tracking is consistent with
species' population crashes (and so extinctions) becoming less
predictable during previous mass extinction events
\supercite{Levinton2001,Jablonski2004}. We call upon others to examine
the additional drivers of our metric of relative change, and to extend
its definition to include more nuanced definitions of climate
space. We predict, however, that if the current mass extinction event
continues, it is likely that patterns of idiosyncrasy among species'
declines and distributional changes will become more common.

@@

Climate data were downloaded from
\citeauthor{Harris2014}\supercite{Harris2014}'s nine global
time-series. We split the data into yearly (averaged) 0.5$^\circ$
cells, from which we estimated each of our nine climate variables for
each species observation using
\emph{raster}\supercite{Hijmans2019}. Plant trait data were taken from
\citeauthor{Wright2004}\supercite{Wright2004} (leaf lifespan, leaf
mass-per-unit-area, leaf Nitrogen mass, and leaf photosynthetic
capacity), mammal body mass data from the Amniote Trait
Database\supercite{Myhrvold2015}, and bird body mass data from
EltonTraits \supercite{Wilman2014}. We were unable to find sufficient
coverage of the other taxa in this study in open-access trait
databases to facilitate further analysis. We used existing global
mammal\supercite{Faurby2015}, bird\supercite{Jetz2012}, amphibian
\supercite{Jetz2018}, plant (`ALLOTB' phylogeny
\supercite{Smith2018}), and reptiles\supercite{Zheng2016} phylogenies,
and in cases where posterior distributions of trees were available we
used a single draw from that distribution. Missing species were added
into the phylogenies using \emph{congeneric.merge} in
\emph{pez}\supercite{Pearse2015}.

>>>>

\usepackage{xspace}
\newcommand{\eg}{\emph{e.g.}\xspace}

@@

The natural environment is changing rapidly. Over the last century,
habitats have become increasingly fragmented and
isolated\supercite{Haddad2015}, the climate has become warmer, and
extreme climate events more frequent\supercite{IPCC2014}. Biological
diversity---the species of plants and animals with which we cohabit
the Earth---has responded predictably. Present-day extinction rates
are estimated to be up to three orders of magnitude greater than
background rates \supercite{Pimm2014} and are projected to increase
further over the next several decades \supercite{MEA2005}. Current
estimates suggest that over one million species may be threatened with
extinction\supercite{IPBESland2019}, and while the main human-induced
drivers of ongoing biodiversity loss are habitat loss
\supercite{Newbold2015} and fragmentation \supercite{Haddad2015},
climate change is likely to become a dominant driver in the next few
decades\supercite{Parmesan2003,Thuiller2007,Urban2015}. Projected
impacts of climate change on biodiversity have attracted much
attention, but the uncertainty around the magnitude (\emph{e.g.},
7.9\%\supercite{Urban2015} vs.\ 37\%\supercite{Thomas2004} of species)
and timing \supercite{Trisos2020} of future extinctions highlights the
inadequacy of current forecasts.  In the face of changing climate
species must either shift in space, to track favourable conditions, or
time, for example, flowering and breeding earlier.
% I've dropped the migration reference here because that's shifting in
% both space and time, and so sort of ruins the either/or structure
Evidence suggests they are doing both
\supercite{Parmesan2003,Parmesan2006,Menzel2006,Chen2011}, but not all
species are responding equally and impacts on species vary across
environments \supercite{Pinsky2019}. Some species are expanding their
distributions, sometimes invasively \supercite{Hellmann2008}, and some
species are flowering later\supercite{Cook2012}; species may also be
becoming less predictable \supercite{Pearse2017phenology},
perhaps as evolved responses to environmental cues break down.  Simply
quantifying species' mean responses ignores the important variability
in responses among species \supercite{Wolkovich2012}. Developing
robust predictions of ecological responses to climate change thus
requires data from diverse species, ecosystems, and climates to
account for this large inter-specific variation.

The velocity of temperature change---the rate a species would have to
shift its range to maintain constant temperature---has been calculated
at an average of 0.42 km/yr \supercite{Loarie2009}. A meta-analysis of
taxonomic groups estimated the average pole-ward shift in species'
distributions has been approximately 1.69 km/yr \supercite{Chen2011},
suggesting that species might be capable of keeping pace with shifting
climate. Yet even this estimate is derived from just 21 studies with
limited taxonomic coverage (361 birds, 9 mammals, and no plants), and
critically any single point estimate does not reflect the large
variation in responses among species or within taxa. Further, there is
emerging evidence that average temperature may be less important in
driving local extinctions than changes in extremes of temperature
\supercite{Roman-Palacios2020} Here we leverage the vast volume of
distributional data stored in digital data
repositories\supercite{url_gbif}, representing over 10,000 species of
plants, fungi, mammals, birds, reptiles, amphibians, and insects, to
provide the first global synthesis of species distributional responses
to climate change. By contrasting historical niche dimensions with
those of the present day, we can characterise the realised biotic
velocity of climate change, which represents the relative degree to
which species track climate across space.

We show that, on average, species' historical (pre-1980) ranges have
warmed significantly, and that species have generally shifted their
distributions to ameliorate this warming, but not sufficiently to keep
pace with climate change. Yet this general pattern masks a remarkably
degree of variability in species responses, and a surprising number of
species' ranges are now even warmer than if their distributions had
remained static. Further, we find that, while species' niche
dimensions are strongly predicted by their evolutionary history and
functional traits, the same is not true of their degree of
tracking. Thus while species' distributions are reasonably
predictable, their changes in distributions are not.

@@

Where $current_\tau$ is a given quantile ($\tau$) of a species'
climatic distribution across its present-day range, $past_\tau$ the
same quantile of a species' range in the past, and $projected_\tau$
the same quantile of a species' past range under present-day climatic
conditions. This index should be calculated separately for
temperature, precipitation, or whatever environmental variable is of
interest. Thus a $track$ of 0 would indicate a species that has
tracked changing climate perfectly, while a $track$ of 1 would be
consistent with a species that had not moved whatsoever to track
changing environmental conditions. Values between 0 and 1 suggest
moderate tracking, while values greater than 1 or less than 0
represent a change in climate relationship greater than if the species
had not moved at all. Uncertainty in either species' past or current
ranges can be accounted for through a bootstrap procedure, whereby the
past and current distributions are re-sampled to standardise sampling
across time periods (see methods).

Our approach has two main innovations: its (1) focus on the whole of a
species' distribution and (2) comparison of observed change with a
null expectation. (1) When estimating species' range distributions we
typically focus on species' limits\supercite{Sexton2009} (\emph{e.g.},
upper and lower quantiles) while in community ecological
\supercite{Tilman2004} and macro-evolutionary \supercite{Zanne2018}
studies we tend to focus on central tendencies (\emph{e.g.}, the
median). However, there is no strong \emph{a priori} reason that
species can, or must, track all quantiles equally. Indeed some
evidence suggests different constraints on species' warm-range limits
versus their cool-range limits, with biotic interactions more
influential in the former, and climate determining the
latter\supercite{Freeman2018}. It is possible, therefore, that the
drivers of species' range edges and centres may be diverging under
climate change. Our approach is to check all quantiles of a species'
distribution. (2) The magnitude of experienced and projected climate
change shows substantial geographical variation
\supercite{Loarie2009}, and so species' responses might thus be
expected to also vary in space. We suggest that species'
distributional responses should be measured on the basis of the
magnitude of change to which they are exposed, and the denominator of
our index re-scales species' change around an expectation under
distributional stasis.  This change reflects the magnitude of
ecological and evolutionary selection, and while there are many ways
of addressing this, we suggest ours is perhaps the simplest and most
transparent, so represents a fair starting point. We note that we
could also rescale our index according to the distance that would need
to be travelled to maintain climate stasis (\emph{sensu}
\citeauthor{Loarie2009}\supercite{Loarie2009}), but prefer to present
here a unitless index of tracking, which is simpler to compare across
taxa and between biogeographic regions.

\textbf{Niche tracking across the tree of life throughout the past
  sixty years.} To assess the extent to which species have tracked
climate, we used the most comprehensively sampled time-series data for
species' distributions (GBIF\supercite{url_gbif}) and climate (UEA's
CRU\supercite{Harris2014}). This represents over 630 million raw
occurrence records collected over the past century. We focus on five
quantiles: the \nth{5}, \nth{25}, \nth{50} (median), \nth{75}, and
\nth{95}. Here, we present species' range data collected from
1955--2015 for XXX species (XXX plants, XXX fungi, XXX mammals, XXX
birds, XXX reptiles and amphibians, and XX insects), each observation
is matched to one of nine climate variables in the year in which they
were sampled. In the methods we give more detail of our reproducible
analysis pipeline, and in the supplementary materials provide
instructions for repeating our analyses with different data curation
and cleaning choices. Our metric was designed to account for known
biases in the GBIF data \supercite{Beck2014}, and in the methods we
outline our simulations which address the influence of sampling
uncertainty.

Almost all species are tracking some climate variable through time
(99.5\% have at least one $track_\tau$ between 0 and 1). Most species
are, to some extent, tracking temperature (see Figure \ref{overview});
65\% track at least one quantile and 43\% track at least
two. Surprisingly, our three temperature indices (mean, minimum, and
maximum temperature) are the least-tracked of any climate variable,
while precipitation is the most-tracked (86\% of species track at
least one of its quantiles). Notably, however, [only XXX\% of/no] taxa
are tracking all quantiles of each climate variable, and the majority
have `overshot' and experienced more change in one aspect of their
niche than if their distributions had remained static (XXX\% have at
least one $track_\tau$ less than 0; for example, becoming colder in a
warming region). Thus while our results confirm that, by-and-large,
species are tracking climate change, the degree of tracking varies
across species, taxa, and aspects of climate (see Figure
\ref{overview}). For example, while fungi and insects are, on the
whole, broadly tracking the median of temperature (median
$track_{50\%}$ of XXX and XXX, respectively), plants and mammals track
temperature (median $track_{50\%}$ of XXX and XXX, respectively) more
poorly than precipitation (median $track_{5\%}$ of XXX and XXX,
respectively).
% WDP: We need to integrate the idea that species are trading off
% different aspects of climate. Probably in the PCA section that comes
% later.

\textbf{Idiosyncratic drivers of species' tracking.} To examine the
potential drivers of species' ranges and range tracking, we first
examined the phylogenetic signal of species' present and past climate
envelopes. We focused on mammals, birds, reptiles, and plants, since
we were able to find broadly inclusive, dated phylogenies for these
taxa. We found strong evidence of phylogenetic signal in species' past
(median Pagel's $\lambda$\supercite{Pagel1999} of XXX) and current
(median $\lambda$ of XXX) distributions, but limited signal in our
index (median $\lambda$ of $<$XXX). Past climate showed stronger
phylogenetic signal than our index in XXX of all 180 (4 taxa, 5
quantiles, and 9 climate variables) possible taxon-climate-quantile
combinations (XXX for current climate; see Figures \ref{phylo} and
\ref{traits-signal}). This confirms the known result that species'
current-day and past ranges are strongly influenced by
evolution\supercite{Peterson1999,Wiens2010}, but suggests different
constraints to contemporary niche tracking.

We additionally examined species' traits as potential drivers of
species' distributions and potential to track changing environment. We
focused on mammals, birds, and plants, since global, openly-released
trait datasets exist for these taxa. Again, we found statistically
significant associations between traits and past distributions of XXX
of 270 (at $\alpha_{5\%}$; 6 traits, 9 climate variables, and 5
quantiles) taxon-climate-quantile combinations (193 for current
distributions), but not between the traits and our tracking index (13
combinations or 5\%; see Figure \ref{traits-signal}). The traits we
used are strongly associated with species' climate
responses\supercite{Wright2004,McCain2014,Estrada2016} (hence their
strong predictive power of species' distributions, but not degree of
tracking), we emphasise that we do not consider species' dispersal
abilities\supercite{Schloss2012}. This is in part because of limited
data, but also because of a potential circularity in predicting
species' observed movements (our tracking index) as a function of
their measured movement ability (from movement). Our findings, while
surprising, are in keeping with emerging evidence that species' traits
are often \supercite{Angert2011,MacLean2017} (but not always
\supercite{Sunday2015,Pacifici2020}) a poor predictor of change at
range edges.

\textbf{Correlations among climate axes force trade-offs in the
  magnitude and kind of tracking species can maintain.} It is well
known that the axes of global climate are not independent and that
they are not changing independently\supercite{Harris2014,IPCC2014},
and so we would na\ively expect species' relative tracking of climate
axes to show similar patterns. As we show in Figure \ref{pca},
principal components analysis of species' current and past climate
distributions, which we refer to as their climate space, shows strong
correlations across climate variables. Yet our index does not:
species' relative degree of tracking is both of much higher dimension
(the amount of variance explained by each principal component axis is
similar) and shows relatively less correlation across climates
axes. Thus while species' positions in climate space show strong
associations, species' relative tracking of their position in climate
space does not. This result suggests that the ecological and
evolutionary rules that determine species' climate relationships are
changing, explaining why species' traits and evolutionary history are
relatively poor predictors of species' niche tracking. That species
are simultaneously tracking some aspects of climate while overshooting
others (see above) results from species trading off different,
correlated aspects of their climatic niche requirements.

There are plausible but unlikely alternative explanations for our
results, of which we highlight three. (1) Our climate variables are
not all-encompassing, and species may be closely tracking other
aspects of climate or environment that we do not consider here (\eg,
geology). Yet we do detect strong pattern in current and past
distributions, such that the absence of pattern in our tracking
indices is deserving of explanation and consideration. (2) Competition
and facilitation among species may mediate species' ability to track
climate\supercite{Davis1999}. Yet, again, they are surely also likely
to have driven species' past distributions, begging the question as to
why species' past and present distributions show such strong trait and
phylogenetic correlates but their relative tracking do not. (3) If
species were perfectly tracking climate, $current_\tau$ would be equal
to $past_\tau$ and so our index values could reflect sampling error or
some other purely stochastic input. If so, our empirical indices
(which are bootstrapped to account for sampling differences) should be
centred at 0, which they are not (see figure \ref{overview}). Further,
as we show in the supplementary materials, we find strong impacts of
non-random change even when holding sampling constant, and non-random
associations and trade-offs in species' niche-tracking.

\clearpage
\subsection*{Conclusion}
Our results are consistent with the processes underlying species'
responses to climate change being distinct from those that generated
their distributions in the past. This is in keeping with emerging
evidence from the paleo-ecological literature that species' historic
ranges were more variable than previously thought, and seemingly
influenced by different environmental factors than today
\supercite{Veloz2012,Maguire2015}. During and after the last ice-age,
for example, `no-analogue' conditions generated previously-unseen
species assemblages \supercite{Williams2007,Veloz2012}. Species simply
cannot track all climate axes of their niches simultaneously, and the
idiosyncrasies among species' different tracking indices likely
reflects trade-offs among niche axes.  Species' climate-change
responses are also inherent context-dependent, and different
ecosystems and habitats have different drivers
\supercite{Bertrand2016,Antao2020}, and biotic interactions can both
mitigate and enhance the impact of climate \supercite{Vandvik2020}.
Equally, our observation of idiosyncrasy in species' degree of
tracking is consistent with species' population crashes (and so
extinctions) becoming less predictable during previous mass extinction
events \supercite{Levinton2001,Jablonski2004}. Perhaps the key to
understanding and predicting these seemingly idiosyncratic differences
in niche tracking will be to understand the distinction between
species' fundamental and realised niches. By focusing on the
underlying cut-offs and boundaries that define species' tolerances and
requirements, we may be able to find pattern among the noise of
species' responses.

@@

Climate data were downloaded from
\citeauthor{Harris2014}\supercite{Harris2014}'s nine global
time-series of mean, minimum, maximum temperature, precipitation,
`rainy-day' frequency, vapour pressure, potential evapotranspiration,
`frost day' frequency, and cloud cover. Given recent evidence of the
divergent roles of mean, minimum, and maximum temperature in driving
local extinctions \supercite{Roman-Palacios2020}, we present track
indices for all three aspects of temperature. We split the data into
yearly (averaged) 0.5$^\circ$ cells, from which we estimated each of
our nine climate variables for each species observation using
\emph{raster}\supercite{Hijmans2019}. Plant trait data were taken from
\citeauthor{Wright2004}\supercite{Wright2004} (leaf lifespan, leaf
mass-per-unit-area, leaf Nitrogen mass, and leaf photosynthetic
capacity), mammal body mass data from the Amniote Trait
Database\supercite{Myhrvold2015}, and bird body mass data from
EltonTraits \supercite{Wilman2014}. We were unable to find sufficient
coverage of the other taxa in this study in open-access trait
databases to facilitate further analysis. We used existing global
mammal\supercite{Faurby2015}, bird\supercite{Jetz2012}, amphibian
\supercite{Jetz2018}, plant (`ALLOTB' phylogeny
\supercite{Smith2018}), and reptiles\supercite{Zheng2016} phylogenies,
and in cases where posterior distributions of trees were available we
used a single draw from that distribution. Missing species were added
into the phylogenies using \emph{congeneric.merge} in
\emph{pez}\supercite{Pearse2015}.
